---
title: "QB Salaries"
author: "Gaurav Law"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggrepel)
library(nflreadr)
library(nflplotR)
library(nflfastR)
library(nflverse)
```


```{r}
nflreadr::clear_cache()
contracts <- nflreadr::load_contracts()
load_teams()
head(contracts, n = 10)
```

```{r}
top_qb_contracts <- contracts %>% 
  filter(position == "QB") %>%
  arrange(year_signed) %>%
  ungroup() %>%
  group_by(otc_id) %>%
  mutate(contract_num = row_number())

last_contracts <- top_qb_contracts %>%
  group_by(otc_id) %>%
  summarize(last_num = max(contract_num))

top_qb_contracts <- top_qb_contracts %>%
  left_join(last_contracts, by = "otc_id") %>%
  filter(contract_num == last_num)
top_qb_contracts

head(top_qb_contracts)

```


```{r}
top_qb_contracts <- top_qb_contracts %>%
  left_join(load_teams(), by = c('team' = 'team_nick')) %>% arrange(-apy)

head(load_teams())

top_qb_contracts[10, "player"] <- "Matthew Stafford"

view(top_qb_contracts)
```


```{r pbp}
pbp <- load_pbp(2018:2022)
pbp_23 <- load_pbp(2023)
```

```{r}
rosters <- load_rosters(2018:2022) %>% distinct(gsis_id, .keep_all = TRUE)
rosters_23 <- load_rosters(2023) %>% distinct(gsis_id, .keep_all = TRUE)

head(load_rosters(2018))

pbp <- pbp %>% left_join(rosters, by = c("id" = "gsis_id"))
pbp_23 <- pbp_23 %>% left_join(rosters_23, by = c("id" = "gsis_id"))

head(pbp)
```

```{r}
qbs <- pbp %>% 
  filter(season_type == "REG", !is.na(epa)) %>%
  filter(pass == 1 | rush == 1) %>%
  group_by(id, full_name) %>%
  summarize(
  q_epa = mean(qb_epa),
  cpoe = mean(cpoe, na.rm = T),
  n_dropbacks = sum(pass),
  n_plays = n(),
  team = last(team)) %>%
  left_join(top_qb_contracts, by = c("full_name" = "player")) %>% filter(n_dropbacks > 100 & n_plays > 500) %>% arrange(-q_epa)

head(qbs, n = 20)
view(qbs)

qb_40 <- head(qbs, n = 40)
qb_40
```


```{r}
qb_40 %>%
  ggplot(aes(x = q_epa, y = apy/1000000)) +
  #horizontal line with mean EPA
  geom_hline(yintercept = mean(qb_40$apy/1000000), color = "red", linetype = "dashed", alpha=0.5) +
  #vertical line with mean CPOE
  geom_vline(xintercept =  mean(qb_40$q_epa), color = "red", linetype = "dashed", alpha=0.5) +
  #add points for the QBs with the logos (this uses nflplotR package)
  geom_nfl_logos(aes(team_abbr = team_abbr), width = qb_40$n_plays / 45000, alpha = 0.75) +
  #add names using ggrepel, which tries to make them not overlap
  geom_text_repel(aes(label=full_name)) +
  #add a smooth line fitting cpoe + epa
  stat_smooth(geom='line', alpha=0.5, se=FALSE, method='lm')+
  #titles and caption
  labs(x = "QB Epa/play",
       y = "Amount per year",
       title = "Quarterback Efficiency vs. Annual Salaries, 2018-2022",
       caption = "Data: @nflfastR\nAuthor: Gaurav Law") +
  theme_bw() +
  #center title
  theme(
    plot.title = element_text(size = 15, hjust = 0.5, face = "bold")
  ) +
  #make ticks look nice
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))

```

```{r}
qbs_current <- pbp_23 %>% 
  filter(season_type == "REG", !is.na(epa)) %>%
  filter(pass == 1 | rush == 1) %>%
  group_by(id, full_name) %>%
  summarize(
  q_epa = mean(qb_epa),
  cpoe = mean(cpoe, na.rm = T),
  n_dropbacks = sum(pass),
  n_plays = n(),
  team = last(team)) %>%
  left_join(top_qb_contracts, by = c("full_name" = "player")) %>% filter(n_dropbacks > 300 & n_plays > 200) %>% arrange(-n_dropbacks)

qbs_current
```

```{r}
qbs_current %>%
  ggplot(aes(x = q_epa, y = apy)) +
  #horizontal line with mean EPA
  geom_hline(yintercept = mean(qbs_current$apy), color = "red", linetype = "dashed", alpha=0.5) +
  #vertical line with mean CPOE
  geom_vline(xintercept =  mean(qbs_current$q_epa), color = "red", linetype = "dashed", alpha=0.5) +
  #add points for the QBs with the logos (this uses nflplotR package)
  geom_nfl_logos(aes(team_abbr = team_abbr), width = qbs_current$n_plays / 12000, alpha = 0.75) +
  #add names using ggrepel, which tries to make them not overlap
  geom_text_repel(aes(label=full_name), max.overlaps = 20) +
  #add a smooth line fitting cpoe + epa
  stat_smooth(geom='line', alpha=0.5, se=FALSE, method='lm')+
  #titles and caption
  labs(x = "QB Epa/play",
       y = "Amount per year",
       title = "Current Quarterback Efficiency vs. Annual Salaries",
       caption = "Data: @nflfastR\nAuthor: Gaurav Law") +
  theme_bw() +
  #center title
  theme(
    plot.title = element_text(size = 15, hjust = 0.5, face = "bold")
  ) +
  #make ticks look nice
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))

```

q_full <- lm(q_epa ~ cpoe*apy + n_dropbacks, data = qbs_current)
q_fitted <- lm(q_epa ~ cpoe*apy, data = qbs_current)
summary(q_fitted)
c(AIC(q_full, q_fitted), BIC(q_full, q_fitted))
```

## A one-percent increase in completion percentage over expected corresponds to a 0.0389-unit increase in QB EPA. This means that there is a very small positive relationship between how good a quarterback is and how often their expectation to complete their passes relative to how often they do. 

## Similarly, a $1M-increase in average salary per year corresponds to a 0.0021-unit increase in QB EPA. This means that salary doesn't really effect how good a quarterback is.

## In addition, the fact that our AIC is small tells us that we have a good model regardless. The model without the number of dropbacks is better than the one with, telling us that dropbacks don't really matter in how good a QB is.



```{r}
p_stats <- load_player_stats()
names(p_stats)
```

Yards Per Route Run

```{r}
pbp_for_routes <- load_participation(2018:2022, include_pbp = TRUE)
players <- load_players()

#extract offensive player IDs

player_rows_pbp <- pbp_for_routes |>
  mutate(offense_player_id = str_extract_all(offense_players, "\\d{2}-\\d{7}")) |>
  unnest(offense_player_id)

#add player info to dataset

player_rows_pbp <- left_join(player_rows_pbp, players, by = c("offense_player_id" = "gsis_id"))

#Example of a 2022 season summary for WRs 
#results include playoff games and targeted plays with penalty yardage
#"routes run" includes passing plays where a player was assigned to block
#WRs almost never block on pass plays, but TEs and RBs sometimes do

routes_run_data <- player_rows_pbp |>
  #add additional season or situational filters here, if desired
  filter(pass == 1, season == 2022,
         position_group == "WR") |>
  group_by(offense_player_id) |>
  summarise(player_name = last(display_name),
            position = last(position_group),
            team = last(posteam),
            routes_run = n(),
            targets = length(offense_player_id[offense_player_id == receiver_id & !is.na(receiver_id)]),
            receiving_yards = sum(yards_gained[offense_player_id == receiver_id & !is.na(receiver_id) & !is.na(yards_gained)]),
            first_downs = sum(first_down[offense_player_id == receiver_id & !is.na(receiver_id) & !is.na(first_down)]),
            touchdowns = sum(touchdown[offense_player_id == receiver_id & !is.na(receiver_id) & !is.na(touchdown)])) |>
  mutate(targets_per_route_run = targets / routes_run,
         yards_per_route_run = receiving_yards / routes_run,
         #This is Timo Riske's formula for adjusted yards per route run
         adj_yprr = (receiving_yards + 9*first_downs + 11*touchdowns) / routes_run) |>
  select(-offense_player_id) |>
  filter(routes_run > 200) |>
  arrange(-yards_per_route_run)
```
